Higher-Order unification for free!

### 2024-04-17 paper/main.tex
abella

\begin{abstract}
  Specifying and implementing a proof system from scratch requires significant effort.
  Logical Frameworks and Higher Order Logic Programming Languages provide
  dedicated, high-level meta languages to facilitate this task in two
  % key 
  ways: 1) variable binding and substitution are for free when meta language binders
  represent object logic ones; 2) proof construction, and %even 
  proof search, are
  greatly simplified by leveraging the unification procedure provided by the meta language.
  Notable examples of meta languages are Twelf~\cite{twelf},
  $\lambda$Prolog~\cite{miller_nadathur_2012}, Beluga~\cite{Beluga} and
  Isabelle~\cite{10.1007/978-3-540-71067-7_7}
  which have been %utilized
  used to implement or specify %various
  many
  formal systems such as
  First Order Logic~\cite{felty88cade},
  Set Theory~\cite{10.1007/BF00881873},
  Higher Order Logic~\cite{books/sp/NipkowPW02}, and %even
  the Calculus of
  Constructions~\cite{felty93lics}.

@@

We considered mechanizing our results with Abella, especially in the early phases.
Unfortunately we could not. One reason
is that eliminating the cut operator, although possible, has a high cost in
verbosity.
% he same for the \elpiIn{name}
%\todo{is it sill described?} predicate.
Another problem is that the logic of Abella does not let one quantify over
predicates and combinators like \elpiIn{forall} are used pervasively in our
code.
\todo{finish}

>>>>

\begin{abstract}
  Specifying and implementing a proof system from scratch requires significant effort.
  Logical Frameworks and Higher Order Logic Programming Languages provide
  dedicated, high-level meta languages to facilitate this task in two
  % key 
  ways: 1) variable binding and substitution are for free when meta language binders
  represent object logic ones; 2) proof construction, and %even 
  proof search, are
  greatly simplified by leveraging the unification procedure provided by the meta language.
  Notable examples of meta languages are Elf~\cite{elf}, Twelf~\cite{twelf},
  $\lambda$Prolog~\cite{miller_nadathur_2012}, Beluga~\cite{Beluga}, Abella~\cite{gacek2008abella} and
  Isabelle~\cite{10.1007/978-3-540-71067-7_7}
  which have been %utilized
  used to implement or specify %various
  many
  formal systems such as
  First Order Logic~\cite{felty88cade},
  Set Theory~\cite{10.1007/BF00881873},
  Higher Order Logic~\cite{books/sp/NipkowPW02}, and %even
  the Calculus of
  Constructions~\cite{felty93lics}.

@@

We considered mechanizing our results with Abella~\cite{gacek2008abella}, especially in the early phases.
Unfortunately we could not. One reason
is that eliminating the cut operator, although possible, has a high cost in
verbosity.
% he same for the \elpiIn{name}
%\todo{is it sill described?} predicate.
Another problem is that the logic of Abella does not let one quantify over
predicates and combinators like \elpiIn{forall} are used pervasively in our
code.
\todo{finish}
